---
title: "CRAM & Macroalgae"
format: html
editor: visual
---

## SoCal Bight Data Analysis - Macroalgae & CRAM Data

|                                                                                                               |
|------------------------------------------------------------------------|
| Question 2: **What estuaries in southern California have the highest vascular plant functional performance?** |
|                                                                                                               |

### Load Packages

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(dplyr)
library(readxl)
```

### Import Macroalgae Files

```{r}
Metadata <-read_excel("macroalgae-2021.xlsx", sheet=1)
algaecover2021 <-read_excel("macroalgae-2021.xlsx", sheet=2)
floating2021 <-read_excel("macroalgae-2021.xlsx", sheet=3)

```

### Cleaning Macroalgae 2021 Data

-   Batiquitos Lagoon

-   Goleta Slough

-   Malibu Lagoon

-   Newport Bay

-   San Elijo Lagoon

-   Ventura River

#### Clean Macroalgae Data

```{r}
#
algaecover21 <- algaecover2021 |> filter(estuaryname == "Malibu Lagoon" | estuaryname == "San Elijo Lagoon" | estuaryname == "Batiquitos Lagoon" | estuaryname == "Goleta Slough" |estuaryname == "Newport Bay" | estuaryname == "Ventura River")

floating21 <- floating2021 |> filter(estuaryname == "Malibu Lagoon" | estuaryname == "San Elijo Lagoon" | estuaryname == "Batiquitos Lagoon" | estuaryname == "Goleta Slough" |estuaryname == "Newport Bay" | estuaryname == "Ventura River")


head(algaecover21)
head(floating21)
```

#### CRAM data

```{r}
CRAM <-read_excel("ecoatlas_cram_update.xlsx", sheet=3)

CRAM
```

#### 

#### 

```{r}

```

#### Bind

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}

```

### 
