---
title: "CRAM & Macroalgae"
format: html
editor: visual
---

## SoCal Bight Data Analysis - Macroalgae & CRAM Data

|                                                                                                               |
|------------------------------------------------------------------------|
| Question 2: **What estuaries in southern California have the highest vascular plant functional performance?** |
|                                                                                                               |

### Load Packages

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(dplyr)
library(readxl)
```

### Import Macroalgae Files

```{r}
Metadata <-read_excel("macroalgae-2021.xlsx", sheet=1)
algaecover2021 <-read_excel("macroalgae-2021.xlsx", sheet=2)
floating2021 <-read_excel("macroalgae-2021.xlsx", sheet=3)

```

### Cleaning Macroalgae 2021 Data

-   Batiquitos Lagoon

-   Goleta Slough

-   Malibu Lagoon

-   Newport Bay

-   San Elijo Lagoon

-   Ventura River

#### Clean Macroalgae Data

```{r}
#
algaecover21 <- algaecover2021 |> filter(estuaryname == "Malibu Lagoon" | estuaryname == "San Elijo Lagoon" | estuaryname == "Batiquitos Lagoon" | estuaryname == "Goleta Slough" |estuaryname == "Newport Bay" | estuaryname == "Ventura River")

floating21 <- floating2021 |> filter(estuaryname == "Malibu Lagoon" | estuaryname == "San Elijo Lagoon" | estuaryname == "Batiquitos Lagoon" | estuaryname == "Goleta Slough" |estuaryname == "Newport Bay" | estuaryname == "Ventura River")



head(floating21)
```

#### CRAM data

```{r}
CRAM1 <-read_excel("ecoatlas_cram_update.xlsx", sheet=2)
CRAM2 <-read_excel("ecoatlas_cram_update.xlsx", sheet=3)

CRAM3 <- CRAM2 |> filter(aaname == "Malibu Lagoon" | aaname == "San Elijo Lagoon" | aaname == "Batiquitos Lagoon" | aaname == "Goleta Slough" |aaname == "Upper Newport Bay" | aaname == "Ventura River" | aaname == "Sweetwater Marsh" | aaname == "Tijuana River Estuary" | aaname == "San Dieguito Lagoon")

CRAM3
```

#### 

#### Preliminary Figure 

```{r}
CRAM_Analysis <- aov(indexscore ~aaname, data=CRAM3)

plottyplot <- ggplot(CRAM3, aes(x = aaname, y = indexscore)) + 
  geom_point() + stat_summary(fun = mean, geom = "line") + labs(x = "Sites") + labs(y = "Cram Index Score") 
```

#### Bind

```{r}


```

### 
