```{r}
library(tidyverse)
library(googlesheets4)
library(readxl)
library(ggtext)
library(ggplot2)
```

```{r}
#import data
read.csv("Combined_Cordgrass.csv")
read.csv("Combined_Vegetation.csv")
```

my_color_vector \<- c("darkblue","blue","lightblue","#C2B280")

SpeciesCount \<- ggplot(df.vegdiverse) +

#dodge allows multiple habitats on one plot

geom_col(aes(y = count,x=Estuary, fill=Habitat),position="dodge")

#adjust agle of estuary names for legibility

print(SpeciesCount)+theme(axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5)) +

scale_fill_manual(values = my_color_vector)+labs(y="Species count",title="Species Diversity by Habitat & Estuary")

```{r}

#create raw data frame
df.veg.raw<-read.csv("Combined_Vegetation.csv")

#finding percent native for each estuary
df.percent<-df.veg.raw|>
count(Estuary, Status) %>%
group_by(Estuary)|>
mutate(percent = n/sum(n)*100) %>%
select(-n) %>%
spread(Status, percent)

#plotting percent native
Native.percentPlot <- ggplot(df.percent) +
aes(y = native,x=Estuary) +
geom_col(aes(y = native,x=Estuary,fill=Estuary))+
  
#Changing the scale since all values above 75%
coord_cartesian(ylim = c(75, 100))+
scale_y_continuous(breaks = seq(75, 100, by = 5))

#Fixing crowded x-axis
print(Native.percentPlot)+theme(axis.text.x = element_text(angle = 30, hjust = 0.5,
vjust = 0.5))

#plotting percent cover for each estuary

Plot2 <- ggplot(df.veg.raw) +
aes(y = VegetatedCover,x=Estuary) +
geom_boxplot()
print(Plot2)
#summary of total count of species at each site
countspec<-df.veg.raw|>
group_by(Estuary)|>
summarise(count = n_distinct(ScientificName))
#Question:
#Do the Estuaries differ in plant species diversity for each habitat type?

df.vegdiverse<-df.veg.raw|>
group_by(Estuary,Habitat)|>
summarise(count = n_distinct(ScientificName)) %>%
  #order habitat types by elevation
mutate(Habitat = factor(Habitat, levels = c("Brackish marsh", "Low marsh", "Mid marsh"
, "High marsh")))|>
  #order Estuaries by size
mutate(Estuary = factor(Estuary, levels = c("Malibu Lagoon", "Ventura River", "Goleta Slough"
, "San Elijo Lagoon", "Batiquitos Lagoon", "Newport Bay")))

#adding custom color
my_color_vector <- c("darkblue","blue","lightblue","#C2B280")
SpeciesCount <- ggplot(df.vegdiverse) +
#dodge allows multiple habitats on one plot
geom_col(aes(y = count,x=Estuary, fill=Habitat),position="dodge")+facet_wrap(~Habitat)
#adjust agle of estuary names for legibility
jpeg("SpeciesCount.jpg", width = 4.5, height = 4, units = 'in', res = 300)
print(SpeciesCount)+theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
scale_fill_manual(values = my_color_vector)+labs(y="Species count",title="Species Diversity by Habitat & Estuary")
dev.off()
SpeciesCount2 <- ggplot(df.vegdiverse) +
#dodge allows multiple habitats on one plot
geom_col(aes(y = count,x=Estuary, fill=Habitat),position="dodge")
#adjust agle of estuary names for legibility
SpeciesCount2<-print(SpeciesCount2)+theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
scale_fill_manual(values = my_color_vector)+labs(y="Species count",title="Species Diversity by Habitat & Estuary")


```
