---
title: "Yvonne EDA"
format: html
editor: visual
---

# Assessment of California Cordgrass and Ridgway's Rail at three estuaries in Southern California

## Load packages

```{r}
library(tidyverse)
library(readxl)
library(recipes)
```

## Import Data

```{r}
#import cordgrass data
Cordgrass_original <- read.csv("Combined_Cordgrass2023.csv")
#import rail data
Rail <- read_excel("Rail.xlsx")
```

## Data Cleaning

```{r}
#remove NA data
Cordgrass <- Cordgrass_original|>
  na.omit(Cordgrass_original)
```

## Cordgrass

```{r}
#get mean and se of percent cover and plant height in 3 sites
Summary <- Cordgrass|>
  group_by(Estuary) |> 
  summarize(count_quadrat = n(),
            mean_pc = mean(VegetatedCover, na.rm = TRUE), 
            se_pc = sd(VegetatedCover)/sqrt(count_quadrat),
            mean_height = mean(PlantHeight, na.rm = TRUE),
            se_height = sd(PlantHeight)/sqrt(count_quadrat)
  )
```

1 Which sites have the highest percent cover of cordgrass? 

```{r}
#bar plot of percent cover
Summary |>
ggplot() +
  aes(x = Estuary, y = mean_pc, fill = Estuary) +  
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = (mean_pc - se_pc), ymax = (mean_pc + se_pc)), width = 0.1) +
  labs(y = "Percent Cover(%)")+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))+
  theme_classic() 
  
```

```{r}
#histogram of percent cover
ggplot(Cordgrass) +
  aes(x = VegetatedCover, fill = Estuary) + 
  geom_histogram(binwidth = 10)+
  facet_wrap(Estuary~.)+
  labs(x = "Percent Cover(%)")+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))
```

```{r}
#one-way anova of percent cover
model_pc <- lm(VegetatedCover ~ Estuary, data = Cordgrass)
#check assumptions
plot(model_pc, 1)
plot(model_pc, 2)
```

```{r}
anova(model_pc)
#significant
```

2 How does the height of cordgrass stands vary across sites? 

```{r}
#bar plot of height
Summary |>
ggplot() +
  aes(x = Estuary, y = mean_height, fill = Estuary) +  
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = (mean_height - se_height), ymax = (mean_height + se_height)), width = 0.1)+
  labs(y = "Plant Height(cm)")+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))+
  theme_classic()
```

```{r}
#histogram of height
ggplot(Cordgrass) +
  aes(x = PlantHeight, fill = Estuary) + 
  geom_histogram(binwidth = 10)+
  facet_wrap(Estuary~.)+
  labs(x = "Plant Height(cm)")+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))
```

```{r}
#one-way anova of height
model_height <- lm(PlantHeight ~ Estuary, data = Cordgrass)
plot(model_height, 1)
plot(model_height, 2)
```

```{r}
anova(model_height)
#significant
```

3 How does the abundance of cordgrass vary across sites?

```{r}
#abundance per quadrat
AbundanceperQuadrat <- Cordgrass|>
  group_by(Estuary, Station, Transect, Quadrat) |>
  summarise(abundance = mean(Abundance))
#mean and se of abundance in 3 sites
AbundanceSummary <- AbundanceperQuadrat|>
  group_by(Estuary)|>
  summarise(count_quadrat = n(),
            mean_abun = mean(abundance),
            se_abun = sd(abundance)/sqrt(count_quadrat))
```

```{r}
#bar plot of abundance
AbundanceSummary |>
ggplot() +
  aes(x = Estuary, y = mean_abun, fill = Estuary) +  
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = (mean_abun - se_abun), ymax = (mean_abun + se_abun)), width = 0.1)+
  labs(y = "Abundance(stems/m2)")+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))+
  theme_classic()
```

```{r}
#histogram of abundance
ggplot(AbundanceperQuadrat) +
  aes(x = abundance, fill = Estuary) + 
  geom_histogram(binwidth = 5)+
  facet_wrap(Estuary~.)+
  labs(x = "Abundance(stems/m2)")+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))
```

```{r}
#one-way anova of abundance
model_abundance <- lm(abundance ~ Estuary, data = AbundanceperQuadrat)
plot(model_abundance, 1)
plot(model_abundance, 2)
```

```{r}
anova(model_abundance)
#significant
```

## Ridgway's Rail

```{r}
#Combine summary with rail data
Summarynocount<- Summary|>
  select(!count_quadrat)
AbundanceSummarynocount <- AbundanceSummary |>
  select(!count_quadrat)
Combined_Summary<- inner_join(Summarynocount, AbundanceSummarynocount) 
Combined_Summary_Rail <- inner_join(Combined_Summary, Rail)

```

1 Is there relevance between cordgrass percent cover and Ridgway's Rail number?

```{r}
#plot of cordgrass percent cover and rail pairs
Combined_Summary_Rail |>
  mutate(Fake = 1) |> #form a group
ggplot() +
  aes(x = Estuary, y = mean_pc, fill = Estuary) +  
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = (mean_pc - se_pc), ymax = (mean_pc + se_pc)), width = 0.1)+
  geom_point(aes(x = Estuary, y = RailPairs), cex = 3) +
  geom_line(aes(group = Fake, x = Estuary, y = RailPairs)) + #fake a line
    scale_y_continuous(
    name = "Percent Cover(%)",
    sec.axis = sec_axis(~.,name="Rail(Pair)")#add a second y axis
  )+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))+
  theme_classic()



   

```

2 Is there relevance between cordgrass percent cover and Ridgway's Rail number?

```{r}
#plot of cordgrass height and rail pairs
Combined_Summary_Rail |>
  mutate(Fake = 1) |> #form a group
ggplot() +
  aes(x = Estuary, y = mean_height, fill = Estuary) +  
  geom_col(width = 0.5) +
  geom_errorbar(aes(ymin = (mean_height - se_height), ymax = (mean_height + se_height)), width = 0.1)+
  geom_point(aes(x = Estuary, y = RailPairs), cex = 3) +
  geom_line(aes(group = Fake, x = Estuary, y = RailPairs)) + #fake a line
    scale_y_continuous(
    name = "Plant Height(cm)",
    sec.axis = sec_axis(~.,name="Rail(Pair)")#add a second y axis
  )+
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))+
  theme_classic()
```

3 Is there relevance between cordgrass abundance and Ridgway's Rail number?

```{r}
#plot of cordgrass abundance and rail pairs
Combined_Summary_Rail |>
  mutate(Fake = 1) |>
ggplot(aes(x=Estuary)) +
  geom_col(aes(y=mean_abun, fill = Estuary), width = 0.5)+
  geom_errorbar(aes(ymin = (mean_abun - se_abun), ymax = (mean_abun + se_abun)), width = 0.1)+ 
  geom_point(aes(y=RailPairs/5), cex = 3) +
  geom_line(aes(group = Fake, y = RailPairs/5))+
  
  scale_y_continuous(
    name = "Abundance(stems/m2)",
    sec.axis = sec_axis(~.*5, name="Rail(Pair)")#add a second y axis
  )+
  
  scale_fill_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))+
  theme_classic()      
```

## PCA

```{r}
#select columns for pca
Cordgrass_clean <- Cordgrass|>
  select(Estuary, Station, Transect, Quadrat, PlantHeightReplicate, Abundance, PlantHeight, VegetatedCover)
```

```{r}
#use recipes package to prepare data for pca
cordgrass_recipe <-
  recipe(~., data = Cordgrass_clean) |> 
  update_role(Estuary, Station, Transect, Quadrat, PlantHeightReplicate, new_role = "id") |>
  step_naomit(all_predictors()) |> 
  step_normalize(all_predictors()) |>
  step_pca(all_predictors(), id = "pca") |>
  prep()
cordgrass_pca <- 
  cordgrass_recipe |>
  tidy(id = "pca") 
```

```{r}
# get pca loadings into wider format
pca_wider <- cordgrass_pca |>
  tidyr::pivot_wider(names_from = component, id_cols = terms)
```

```{r}
# define arrow style
arrow_style <- arrow(length = unit(.05, "inches"),
                     type = "closed")
#plot PC1 and PC2
pca_plot1 <-
  juice(cordgrass_recipe) |>
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = Estuary, shape = Estuary), 
             alpha = 0.8, 
             size = 2) +
  scale_colour_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))
 

pca_plot1 +
  geom_segment(data = pca_wider,
               aes(xend = PC1, yend = PC2), 
               x = 0, 
               y = 0, 
               arrow = arrow_style) + 
  geom_text(data = pca_wider,
            aes(x = PC1, y = PC2, label = terms), 
            hjust = 0, 
            vjust = 1,
            size = 5) 
```

```{r}
#plot PC2 and PC3
pca_plot2 <-
  juice(cordgrass_recipe) |>
  ggplot(aes(PC2, PC3)) +
  geom_point(aes(color = Estuary, shape = Estuary), 
             alpha = 0.8, 
             size = 2) +
  scale_colour_manual(values = c('palegreen3','lightskyblue','mediumpurple1'))

pca_plot2 +
  geom_segment(data = pca_wider,
               aes(xend = PC2, yend = PC3), 
               x = 0, 
               y = 0, 
               arrow = arrow_style) + 
  geom_text(data = pca_wider,
            aes(x = PC2, y = PC3, label = terms), 
            hjust = 0, 
            vjust = 1,
            size = 5) 
```
